---
title: "EnhancerInsight prioritization report"
output:
  html_document:
    toc: true
    toc_depth: 3
    number_sections: true
params:
  report_dir: "report"
---

<script src="https://unpkg.com/cytoscape/dist/cytoscape.min.js"></script>

<style>
.cy-report-wrapper {
  display: flex;
  gap: 14px;
  width: 100%;
  min-height: 620px;
  margin: 12px 0 24px 0;
}
.cy-report-network {
  flex: 1 1 76%;
  min-width: 0;
  height: 620px;
  border: 1px solid #d8d8d8;
  border-radius: 6px;
  background: #ffffff;
}
.cy-report-side {
  flex: 0 0 24%;
  min-width: 230px;
  max-height: 620px;
  overflow-y: auto;
  border: 1px solid #d8d8d8;
  border-radius: 6px;
  padding: 10px 12px;
  background: #fafafa;
  overflow-wrap: anywhere;
  word-break: break-word;
}
.cy-report-toolbar {
  margin: 6px 0 8px 0;
}
.cy-report-toolbar button {
  margin-right: 6px;
  padding: 4px 10px;
}
.cy-report-legend {
  font-size: 0.92em;
  margin: 4px 0 8px 0;
}
.cy-dot {
  display: inline-block;
  width: 11px;
  height: 11px;
  border: 1px solid #333;
  border-radius: 50%;
  margin: 0 4px 0 10px;
  vertical-align: middle;
}
.cy-dot:first-child {
  margin-left: 0;
}
.cy-focal { background: #e53935; }
.cy-seed { background: #1e88e5; }
.cy-recipient { background: #cfd3d7; }

@media (max-width: 900px) {
  .cy-report-wrapper {
    display: block;
  }
  .cy-report-network {
    width: 100%;
    height: 540px;
  }
  .cy-report-side {
    width: auto;
    max-height: none;
    margin-top: 10px;
  }
}

.cy-line {display:inline-block;width:34px;height:0;margin:0 5px 2px 10px;vertical-align:middle;}
.cy-line-direct-seed {border-top:3px solid #1e88e5;}
.cy-line-indirect-seed {border-top:3px dashed #1e88e5;}
.cy-line-recipient {border-top:3px solid #7a7a7a;}

.gwas-more {
  display: inline-block;
  margin-top: 4px;
  font-weight: 600;
  color: #555555;
}

</style>


<style>
table td,
table th {
  vertical-align: top !important;
}

/* By default, keep node names, regions, ranks, and numeric values on one line. */
table td {
  white-space: nowrap !important;
  word-break: normal !important;
  overflow-wrap: normal !important;
}

/* Only long GWAS annotation columns are allowed to wrap. */
table td.gwas-wrap,
table th.gwas-wrap {
  white-space: normal !important;
  word-break: break-word !important;
  overflow-wrap: anywhere !important;
  vertical-align: top !important;
}

/* Keep the first column readable and prevent chromosome coordinates from
   breaking in the middle. */
table td:first-child,
table th:first-child {
  white-space: nowrap !important;
  min-width: 120px;
}
</style>


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE,message=FALSE,warning=FALSE,fig.width=9,fig.height=5.5,dpi=150)
source("R/io.R")
source("R/utils.R")
source("R/network_utils.R")
report_dir <- resolve_report_dir(params$report_dir)
tables <- load_tables(report_dir)
mode <- detect_mode(tables)
score_col <- final_score_col(mode, tables$ranking)
```

```{r child-overview, child='sections/01_overview.Rmd'}
```

```{r child-prioritized, child='sections/02_prioritized_candidates.Rmd'}
```

```{r child-rank-change, child='sections/03_rank_changes.Rmd'}
```

```{r child-network, child='sections/04_network_interpretation.Rmd'}
```

```{r child-appendix, child='sections/05_appendix.Rmd'}
```
