---
title: "annoEnhancer Report"
date: "`r format(Sys.Date(), '%m/%d/%Y')`"
output:
  rmdformats::readthedown:
    toc_depth: 3
    number_sections: true
params:
  outdir: NULL        # directory containing: <proj>.annotation.txt, <proj>.m_tf.txt, <proj>.m_SNP.txt, <proj>.m_eQTL.txt, <proj>.target.txt
  proj: NULL          # project prefix, e.g. "K562-CTCF"
  genome: "hg19"      # "hg19" or "hg38"
  refdir: NULL        # directory containing summary tables: TFBS-summary-<genome>.txt, SNP-summary-<genome>.txt, eqtl-summary-<genome>.txt
  tissue_map_nascent: NULL    # optional (2-col): sample <tab> tissue
  tissue_map_sc: NULL    # optional (2-col): sample <tab> tissue
  enhancer_count_file: NULL    # optional (2-col): sample <tab> tissue
  tfbs_summary: NULL  # optional override for TFBS summary file
---

```{css, echo=FALSE}
#content{
    max-width:1920px;
}
```

<style type="text/css">
  div.datatables { height: auto !important;}
  div.main-container { max-width: 100% !important; margin: auto; }
</style>

```{r setup_options, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  results = "asis",
  fig.width = 7,
  fig.height = 7,
  tidy = TRUE,
  tidy.opts = list(comment = FALSE, width.cutoff = 120)
)
options(knitr.graphics.rel_path = FALSE)
```

```{r load_packages, include=FALSE}
suppressPackageStartupMessages({
  library(DT)
  library(dplyr)
  library(ggplot2)
  library(VennDiagram)
  library(grid)
  library(pheatmap)
  library(RColorBrewer)
  library(gplots)
  library(data.table)
})

# WebGestaltR is optional (many HPC envs don't have it). We'll load later if available.
```

```{r helpers, include=FALSE}
normalize_colnames <- function(x) {
  # Make headers robust: convert '-' and spaces to '.'
  x <- gsub(" ", ".", x, fixed = TRUE)
  x <- gsub("-", ".", x, fixed = TRUE)
  x
}

pick_col <- function(df, candidates) {
  cn <- colnames(df)
  for (c in candidates) if (c %in% cn) return(c)
  return(NULL)
}

display_venn_safe <- function(x, ...) {
  # VennDiagram errors when any set is empty; guard and skip instead of failing the report.
  sizes <- sapply(x, length)
  if (any(is.na(sizes)) || any(sizes == 0)) {
    cat("Venn diagram skipped (one or more sets are empty).\n")
    return(invisible(NULL))
  }
  grid.newpage()
  venn_object <- VennDiagram::venn.diagram(x, filename = NULL, ...)
  grid::grid.draw(venn_object)
}

read_two_col_map <- function(path) {
  if (is.null(path) || path == "" || !file.exists(path)) return(NULL)
  m <- read.table(path, sep="\t", header=FALSE, quote="", as.is=TRUE)
  if (ncol(m) < 2) return(NULL)
  m <- m[m$V2 != "", , drop=FALSE]
  if (nrow(m) == 0) return(NULL)
  m
}
```

```{r resolve_inputs, include=FALSE}
# -----------------------
# Parameters
# -----------------------
outdir <- params$outdir
proj <- params$proj
genome <- params$genome
refdir <- params$refdir
tissue_map_nascent <- params$tissue_map_nascent
tissue_map_sc <- params$tissue_map_sc

tfbs_summary <- params$tfbs_summary

if (is.null(outdir) || is.null(proj) || outdir == "" || proj == "") {
  stop("Missing required params: outdir and proj")
}
if (is.null(refdir) || refdir == "") {
  stop("Missing required param: refdir (e.g., ../hg19 or ../hg38)")
}

# Normalize trailing slashes
if (!grepl("/$", outdir)) outdir <- paste0(outdir, "/")
if (!grepl("/$", refdir)) refdir <- paste0(refdir, "/")

# -----------------------
# Resolve key input files
# -----------------------
anno_path   <- paste0(outdir, proj, ".annotation.txt")
m_tf_path   <- paste0(outdir, proj, ".m_tf.txt")
m_snp_path  <- paste0(outdir, proj, ".m_SNP.txt")
m_eqtl_path <- paste0(outdir, proj, ".m_eQTL.txt")
gene_path   <- paste0(outdir, proj, ".target.txt")

# Manhattan input (generated by your additional step)
activity_file  <- paste0(outdir, proj, "-HACER.txt")
regions_file <- NA  # keep NA: you already pre-filtered to GM12878 p300 binding-site regions
#out_z_mat  <- "/Users/cqs_jing/OneDrive - VUMC/annoEnhancer/manuscript/metric_z_region_by_cellline.tsv"
#out_metrics_table <- "/Users/cqs_jing/OneDrive - VUMC/annoEnhancer/manuscript/cellline_origin_metrics.tsv"
accessibility_file <- paste0(outdir, proj, "-scATAC.txt")
enhancer_count_file <- params$enhancer_count_file

# Summary tables in refdir (allow override for TFBS)
if (is.null(tfbs_summary) || tfbs_summary == "") {
  tfbs_summary <- paste0(refdir, "TFBS-summary-", genome, ".txt")
}
snp_summary  <- paste0(refdir, "SNP-summary-", genome, ".txt")
eqtl_summary <- paste0(refdir, "eqtl-summary-", genome, ".txt")

# Tissue map (optional) used by the HACER activity specificity Manhattan plot.
# If not provided, try refdir/HACER-tissue.txt (e.g. hg19/HACER-tissue.txt).
if (is.null(tissue_map_nascent) || tissue_map_nascent == "") {
  tissue_guess <- paste0(refdir, "HACER-tissue.txt")
  if (file.exists(tissue_guess)) {
    tissue_map_nascent <- tissue_guess
  }
}

# Tissue map (optional) used by the single cell type specificity Manhattan plot.
# If not provided, try refdir/dbscATAC-human-tissue-map.txt (e.g. hg19/dbscATAC-human-tissue-map.txt.
if (is.null(tissue_map_sc) || tissue_map_sc == "") {
  tissue_guess_sc <- paste0(refdir, "dbscATAC-human-tissue-map.txt")
  if (file.exists(tissue_guess_sc)) {
    tissue_map_sc <- tissue_guess_sc
  }
}

# Enhancer count map (optional) used by the single cell type specificity Manhattan plot.
# If not provided, try refdir/dbscATAC-enhancer-count.txt(e.g. hg19/dbscATAC-enhancer-count.txt.
if (is.null(enhancer_count_file) || enhancer_count_file == "") {
  count_guess <- paste0(refdir, "dbscATAC-enhancer-count.txt")
  if (file.exists(count_guess)) {
    enhancer_count_file <- count_guess
  }
}


# Sanity checks (do not stop hard except annotation)
if (!file.exists(anno_path)) stop(paste0("Missing annotation file: ", anno_path))
```

```{r load_annotation, include=FALSE}
file <- read.table(anno_path, sep="\t", header=TRUE, quote="", as.is=TRUE, check.names=FALSE)
colnames(file) <- normalize_colnames(colnames(file))

# Function column compatibility
fn_col <- pick_col(file, c("Function.NCBI_RefSeq", "Function.NCBI.RefSeq", "Function.NCBI_RefSeq."))
if (is.null(fn_col)) {
  stop("Cannot find Function column: expected 'Function-NCBI_RefSeq' or 'Function.NCBI_RefSeq'")
}
genomeFeature <- table(file[[fn_col]])

t <- nrow(file)
n <- nrow(file)

# Precompute "all regions"
all_regions <- apply(file[, c(1:3)], 1, paste, collapse="-")
```

<br>

# Summary

## Overall summary

```{r overall_summary_table}
total <- sum(genomeFeature)
genomeFeature["Total"] <- sum(genomeFeature)

genomeFeature_df <- as.data.frame(as.matrix(genomeFeature))
genomeFeature_df$Fraction <- genomeFeature_df[,1] / total
genomeFeature_df <- data.frame(
  `Genome Feature` = rownames(genomeFeature_df),
  Number = genomeFeature_df[,1],
  Fraction = genomeFeature_df$Fraction,
  row.names = NULL
)

DT::datatable(
  genomeFeature_df,
  rownames = FALSE,
  extensions = "Buttons",
  options = list(dom = "Bfrtip", buttons = c("excel", "csv"))
) %>% DT::formatRound(columns = c("Number", "Fraction"), digits = 5)
```

<br>

## Summary of support by public super enhancer databases

```{r superenhancer_support, fig.show="hold", out.width="40%"}
# After normalization:
# superEnhancer-SEdb2.0 -> superEnhancer.SEdb2.0
# superEnhancer-dbSUPER -> superEnhancer.dbSUPER
se_se_col <- pick_col(file, c("superEnhancer.SEdb2.0"))
se_db_col <- pick_col(file, c("superEnhancer.dbSUPER"))


  s1 <- apply(file[file[[se_se_col]] == "Y", c(1:3)], 1, paste, collapse="-")
  s2 <- apply(file[!(is.na(file[[se_db_col]])), c(1:3)], 1, paste, collapse="-")

  s <- data.frame(
    Database = c("SEdb", "dbSUPER"),
    Number = c(length(s1), length(s2)),
    `Jaccard Index` = c(
      length(s1) / (t + 1160165 - length(s1)),
      length(s2) / (t + 68894 - length(s2))
    )
  )

  DT::datatable(
      s,
      rownames = FALSE,
      extensions = "Buttons",
      options = list(dom = "Bfrtip", buttons = c("excel", "csv"))
  ) %>% DT::formatRound(columns = 3, digits = 5)

  x <- list(SEdb = s1, dbSUPER = s2)
  display_venn_safe(x, fill = c("#999999", "#E69F00"), cat.fontface = "bold")

```

<br>

## Summary of support by public enhancer databases

```{r enhancer_support, fig.show="hold", out.width="50%"}
# After normalization:
# Enhancer-HACER -> Enhancer.HACER, etc.
e_cols <- c(
  HACER = "Enhancer.HACER",
  VISTA = "Enhancer.VISTA",
  RAEdb = "Enhancer.RAEdb",
  EnsemblRegulatoryBuild = "Enhancer.EnsemblRegulatoryBuild",
  ENCODEBroad.ChromHMM = "Enhancer.ENCODEBroad.ChromHMM",
  NIH.ROADMAP = "Enhancer.NIH.ROADMAP"
)

get_regions_non_na <- function(colname) {
  if (!(colname %in% colnames(file))) return(character(0))
  apply(file[!(is.na(file[[colname]])), c(1:3)], 1, paste, collapse="-")
}

e1 <- get_regions_non_na(e_cols[["HACER"]])
e2 <- get_regions_non_na(e_cols[["VISTA"]])
e3 <- get_regions_non_na(e_cols[["RAEdb"]])
e4 <- get_regions_non_na(e_cols[["EnsemblRegulatoryBuild"]])
e5 <- get_regions_non_na(e_cols[["ENCODEBroad.ChromHMM"]])
e6 <- get_regions_non_na(e_cols[["NIH.ROADMAP"]])

e_tbl <- data.frame(
  Database = c("HACER", "RAEdb", "EnsemblRegulatoryBuild", "ENCODEBroad.ChromHMM", "NIH.ROADMAP"),
  Number = c(length(e1), length(e3), length(e4), length(e5), length(e6)),
  `Jaccard Index` = c(
    length(e1) / (t + 1685280 - length(e1)),
    length(e3) / (t + 436473 - length(e3)),
    length(e4) / (t + 163595 - length(e4)),
    length(e5) / (t + 163330 - length(e5)),
    length(e6) / (t + 9379875 - length(e6))
  )
)

DT::datatable(
  e_tbl,
  rownames = FALSE,
  extensions = "Buttons",
  options = list(dom = "Bfrtip", buttons = c("excel", "csv"))
) %>% DT::formatRound(columns = 3, digits = 5)

x <- list(HACER=e1, RAEdb=e3, EnsemblRegulatoryBuild=e4, ENCODEBroad.ChromHMM=e5, NIH.ROADMAP=e6)
display_venn_safe(
  x,
  fill = c("#999999", "#E69F00", "#56B4E9", "#009E73", "#FB766D"),
  cat.cex = 1,
  cat.fontface = "bold",
  cat.default.pos = "outer",
  cat.dist = c(0.2, 0.15, 0.2, 0.2, 0.15)
)

# Support distribution pie
xx <- data.frame(region=c(e1,e3,e4,e5,e6), num=1)
xx$sup <- ave(xx$num, xx$region, FUN = sum)
xx <- unique(xx)

nsup <- setdiff(all_regions, xx$region)
nsup <- data.frame(region=nsup, sup=0)

xx2 <- rbind(xx[, c("region","sup")], nsup)
support <- table(xx2$sup)
lbls <- paste(names(support), support, sep=":")
pie(support, labels=lbls, col=rainbow(length(support)),
    main="# of support enhancer databases : # of regions")
```

<br>

## Overlap with experiment validated enhancers in VISTA

```{r vista_overlap, fig.show="hold", out.width="20%"}
if (length(e2) == 0) {
  cat("VISTA set is empty or column missing; skipping VISTA overlaps.\n")
} else {
  display_venn_safe(list(HACER=e1, VISTA=e2), cat.fontface="bold", scaled=FALSE,
                    fill=c("#999999","#FF61CC"), cex=2, cat.cex=1.5)
  display_venn_safe(list(RAEdb=e3, VISTA=e2), cat.fontface="bold", scaled=FALSE,
                    fill=c("#E69F00","#FF61CC"), cex=2, cat.cex=1.5)
  display_venn_safe(list(EnsemblRegulatoryBuild=e4, VISTA=e2), cat.fontface="bold", scaled=FALSE,
                    fill=c("#56B4E9","#FF61CC"), cex=2, cat.cex=1.5)
  display_venn_safe(list(ENCODEBroad.ChromHMM=e5, VISTA=e2), cat.fontface="bold", scaled=FALSE,
                    fill=c("#009E73","#FF61CC"), cex=2, cat.cex=1.5)
  display_venn_safe(list(NIH.ROADMAP=e6, VISTA=e2), cat.fontface="bold", scaled=FALSE,
                    fill=c("#FB766D","#FF61CC"), cex=2, cat.cex=1.5)
}
```

<br>

# Activity specificity

```{r activity_specificity_manhattan, fig.height=9, fig.width=14}
if (!file.exists(activity_file)) {
  cat("HACER activity file not found: ", activity_file, "\n")
  cat("Skipping Activity specificity section.\n")
} else {
  # -----------------------
  # Knobs (core)
  # -----------------------
  min_celllines_per_tissue <- 1
  high_quantile <- 0.90
  top_label_n <- 5
  bar_top_n <- 40
  
  # -----------------------
  # Advanced toggles
  # -----------------------
  opt_block_bg      <- TRUE
  opt_top_threshold <- TRUE
  opt_bootstrap_ci  <- FALSE
  
  opt_drop_allzero_regions      <- TRUE
  opt_repel_labels              <- TRUE
  opt_consistent_tissue_palette <- TRUE
  opt_plot_only_big_tissues     <- FALSE
  plot_min_celllines_per_tissue <- 3
  
  opt_intensity_baseline_adjust <- FALSE
  coverage_mode <- "per_cellline"   # "global" | "per_cellline" | "per_tissue" | "global_or_tissue"
  
  threshold_type <- "meanZ_q95"     # "meanZ_q95" | "top_meanZ_median" | "none"

  # -----------------------
  # Helpers
  # -----------------------
  zscore <- function(x) {
    s <- sd(x, na.rm=TRUE)
    if (!is.finite(s) || s == 0) return(rep(0, length(x)))
    (x - mean(x, na.rm=TRUE)) / s
  }

  make_tissue_palette <- function(levels_vec) {
    k <- length(levels_vec)
    if (requireNamespace("Polychrome", quietly=TRUE)) {
      cols <- Polychrome::palette36.colors
      if (k <= length(cols)) return(setNames(cols[seq_len(k)], levels_vec))
    }
    if (requireNamespace("randomcoloR", quietly=TRUE)) {
      cols <- randomcoloR::distinctColorPalette(k)
      return(setNames(cols, levels_vec))
    }
    base <- c(brewer.pal(9,"Set1"), brewer.pal(8,"Dark2"), brewer.pal(8,"Set2"))
    setNames(rep(base, length.out=k), levels_vec)
  }
  
  make_rank_bar <- function(df, value_col, title, higher_better=TRUE, top_n=40) {
    d <- copy(df)
    if (higher_better) d <- d[order(-get(value_col))] else d <- d[order(get(value_col))]
    d <- d[1:min(top_n, nrow(d))]
    d[, cellline := factor(cellline, levels = rev(cellline))]
    ggplot(d, aes(x=cellline, y=.data[[value_col]], fill=tissue_plot)) +
      geom_col(width=0.85) + coord_flip() +
      labs(x=NULL, y=title) +
      theme_classic(base_size=11) +
      theme(panel.grid.major.y = element_blank(),
            panel.grid.minor = element_blank(),
            legend.position="right")
  }

  # -----------------------
  # 1) Read activity; collapse duplicates by MAX
  # -----------------------
  dt_all <- fread(activity_file, header=FALSE, sep="\t",
                  col.names=c("region","cellline","activity"))
  dt_all <- dt_all[, .(activity=max(activity, na.rm=TRUE)), by=.(region, cellline)]
  
  cell <- fread(tissue_map_nascent, header=FALSE, sep="\t", col.names=c("cellline","tissue"))
  cell <- cell[!is.na(cellline) & cellline!="" & !is.na(tissue) & tissue!=""]
  
  # restrict to mapped cell lines
  dt_all <- dt_all[cell, on="cellline", nomatch=0]
  
  # keep tissues with >= N cell lines
  tab_tis <- table(cell$tissue)
  keep_tis <- names(tab_tis[tab_tis >= min_celllines_per_tissue])
  cell <- cell[tissue %in% keep_tis]
  dt_all <- dt_all[cell, on="cellline", nomatch=0]
  
  # optional: restrict to input regions file (usually NA in your case)
  dt <- copy(dt_all)
  if (!is.na(regions_file)) {
    rr <- fread(regions_file, header=FALSE)
    input_regions <- unique(rr[[1]])
    dt <- dt[region %in% input_regions]
#    message("Using regions_file: N=", length(input_regions),
#            " | found in table: N=", length(unique(dt$region)))
  } else {
#    message("No regions_file: using regions present in activity_file: N=", length(unique(dt$region)))
  }
  
  # -----------------------
  # 2) Pseudocount + log transform
  # -----------------------
  pc <- suppressWarnings(min(dt_all$activity[dt_all$activity > 0]))
  if (!is.finite(pc) || pc <= 0) pc <- 1e-12
  
  # -----------------------
  # 3) Build FULL matrix region x cellline (fill missing as 0)
  #    IMPORTANT FIX: compute Intensity/Coverage on this full matrix
  # -----------------------
  mm_dt <- dcast(dt, region ~ cellline, value.var="activity", fill=0)
  regions_u <- mm_dt$region
  mm_dt[, region := NULL]
  mm <- as.matrix(mm_dt)
  rownames(mm) <- regions_u
  rm(mm_dt)
  
  cell_u <- unique(cell[, .(cellline, tissue)])
  cell_u <- cell_u[cellline %in% colnames(mm)]
  setorder(cell_u, tissue, cellline)
  mm <- mm[, cell_u$cellline, drop=FALSE]
  
  # drop all-zero regions across included cell lines
  if (opt_drop_allzero_regions) {
    keep_r <- which(rowSums(mm > 0) > 0)
    mm <- mm[keep_r, , drop=FALSE]
#    message("Dropped all-zero regions: kept ", nrow(mm), " regions.")
  }
  # per cell line: how many regions have non-zero activity
  n_active_regions <- colSums(mm > 0, na.rm = TRUE)
  logM <- log2(mm + pc)
  
  # -----------------------
  # 4) Baseline (optional): compute on ALL regions in dt_all but as FULL matrix too
  #    so baseline isn't affected by sparsity
  # -----------------------
  BaselineIntensity <- NULL
  if (opt_intensity_baseline_adjust) {
    mm_all_dt <- dcast(dt_all, region ~ cellline, value.var="activity", fill=0)
    mm_all_dt[, region := NULL]
    mm_all <- as.matrix(mm_all_dt)
    mm_all <- mm_all[, intersect(colnames(mm_all), cell_u$cellline), drop=FALSE]
    log_all <- log2(mm_all + pc)
    BaselineIntensity <- apply(log_all, 2, median, na.rm=TRUE)
    rm(mm_all_dt, mm_all, log_all)
  }
  
  # -----------------------
  # 5) Coverage thresholds (computed on dt_all full-matrix distribution)
  # -----------------------
  # global threshold from all values (full matrix)
  mm_all_for_thr <- dcast(dt_all, region ~ cellline, value.var="activity", fill=0)
  mm_all_for_thr[, region := NULL]
  log_all_vals <- as.vector(log2(as.matrix(mm_all_for_thr) + pc))
  thr_high_global <- as.numeric(quantile(log_all_vals, high_quantile, na.rm=TRUE))
  rm(mm_all_for_thr, log_all_vals)
  
  # per-cellline thresholds (from full matrix)
  thr_cell <- NULL
  if (coverage_mode == "per_cellline") {
    mm_all2 <- dcast(dt_all, region ~ cellline, value.var="activity", fill=0)
    regs2 <- mm_all2$region; mm_all2[, region := NULL]
    log_all2 <- log2(as.matrix(mm_all2) + pc)
    colnames(log_all2) <- colnames(mm_all2)
    thr_cell <- apply(log_all2, 2, function(x) as.numeric(quantile(x, high_quantile, na.rm=TRUE)))
    rm(mm_all2, log_all2, regs2)
  }
  
  # per-tissue thresholds (from full matrix): pool all celllines within tissue
  thr_tis <- NULL
  if (coverage_mode %in% c("per_tissue", "global_or_tissue")) {
    mm_all3 <- dcast(dt_all, region ~ cellline, value.var="activity", fill=0)
    mm_all3[, region := NULL]
    log_all3 <- log2(as.matrix(mm_all3) + pc)
    colnames(log_all3) <- colnames(mm_all3)
    # map col->tissue
    col_tissue_map <- cell_u[match(colnames(log_all3), cell_u$cellline), tissue]
    thr_tis <- sapply(unique(col_tissue_map), function(tis) {
      cols <- which(col_tissue_map == tis)
      as.numeric(quantile(as.vector(log_all3[, cols, drop=FALSE]), high_quantile, na.rm=TRUE))
    })
    rm(mm_all3, log_all3, col_tissue_map)
  }
  
  # -----------------------
  # 6) Metrics computed on FULL matrix logM (NOT dt long table)
  # -----------------------
  Intensity <- apply(logM, 2, median, na.rm=TRUE)
  #Intensity_q <- 0.75  # for K562, changed to quantile instead of median. 0.75 推荐；更“看强信号”可用 0.90
  #Intensity <- apply(logM, 2, function(x) as.numeric(quantile(x, probs = Intensity_q, na.rm = TRUE)))
  
  if (coverage_mode == "global") {
    Coverage <- apply(logM, 2, function(x) mean(x >= thr_high_global, na.rm=TRUE))
  } else if (coverage_mode == "per_cellline") {
    Coverage <- mapply(function(col, thr) mean(col >= thr, na.rm=TRUE),
                       as.data.frame(logM), thr_cell[colnames(logM)])
    Coverage <- as.numeric(Coverage); names(Coverage) <- colnames(logM)
  } else if (coverage_mode == "per_tissue") {
    Coverage <- sapply(colnames(logM), function(cl) {
      tis <- cell_u[cellline == cl, tissue][1]
      mean(logM[, cl] >= thr_tis[[tis]], na.rm=TRUE)
    })
  } else if (coverage_mode == "global_or_tissue") {
    Coverage <- sapply(colnames(logM), function(cl) {
      tis <- cell_u[cellline == cl, tissue][1]
      thr <- max(thr_high_global, thr_tis[[tis]])
      mean(logM[, cl] >= thr, na.rm=TRUE)
    })
  } else stop("Unknown coverage_mode: ", coverage_mode)
  
  # MeanRank (lower better)
  rankM <- t(apply(logM, 1, function(x) rank(-x, ties.method="average")))
  MeanRank <- colMeans(rankM, na.rm=TRUE)
  #rm(rankM)
  
  # SumZ / MeanZ (higher better)
  rowMean <- rowMeans(logM, na.rm=TRUE)
  rowSd <- apply(logM, 1, sd, na.rm=TRUE)
  rowSd[rowSd == 0 | !is.finite(rowSd)] <- NA_real_
  zM <- (logM - rowMean) / rowSd
  zM[!is.finite(zM)] <- NA_real_
  SumZ <- colSums(zM, na.rm=TRUE)
  MeanZ <- SumZ / nrow(zM)
  #rm(rankM, zM, rowMean, rowSd, mm, logM)
  write_mat_tsv <- function(mat, file) {
    dt_out <- as.data.table(mat, keep.rownames = "region")
    fwrite(dt_out, file = file, sep = "\t", quote = FALSE)
  }
#  write_mat_tsv(zM, out_z_mat)
#  message("Saved: ", out_z_mat)
  
  
  # assemble table
  metrics <- data.table(
    cellline = cell_u$cellline,
    tissue   = cell_u$tissue
  )
  metrics[, n_active_regions := as.integer(n_active_regions[cellline])]
  metrics[, frac_active_regions := n_active_regions / nrow(mm)]
  metrics[, n_regions_total := nrow(mm)]
  metrics[, Intensity := as.numeric(Intensity[cellline])]
  metrics[, Coverage  := as.numeric(Coverage[cellline])]
  metrics[, MeanRank  := as.numeric(MeanRank[cellline])]
  metrics[, SumZ      := as.numeric(SumZ[cellline])]
  metrics[, MeanZ     := as.numeric(MeanZ[cellline])]
  metrics[, n_regions := nrow(mm_u <- dcast(dt, region ~ cellline, value.var="activity", fill=0))]  # same for all; quick label
  #rm(mm_u)
  rm(rankM, zM, rowMean, rowSd, mm, logM)
  
  if (opt_intensity_baseline_adjust) {
    metrics[, BaselineIntensity := as.numeric(BaselineIntensity[cellline])]
    metrics[, Intensity_adj := Intensity - BaselineIntensity]
    metrics[, Intensity_used := Intensity_adj]
  } else {
    metrics[, BaselineIntensity := NA_real_]
    metrics[, Intensity_adj := NA_real_]
    metrics[, Intensity_used := Intensity]
  }
  
  metrics[, Coverage_used := Coverage]
  
  # OriginScore
  metrics[, ZI := zscore(Intensity_used)]
  metrics[, ZC := zscore(Coverage_used)]
  metrics[, ZR := zscore(-MeanRank)]
  metrics[, ZZ := zscore(SumZ)]
  metrics[, OriginScore := ZI + ZC + ZR + ZZ]
#  fwrite(metrics[order(-MeanZ)], file = out_metrics_table, sep = "\t", quote = FALSE)
#  message("Saved metrics table: ", out_metrics_table)
  
  # -----------------------
  # 7) Plot-only tissue grouping
  # -----------------------
  metrics[, tissue_plot := as.character(tissue)]
  if (opt_plot_only_big_tissues) {
    tc <- metrics[, .N, by=tissue_plot]
    big <- tc[N >= plot_min_celllines_per_tissue, tissue_plot]
    metrics[!tissue_plot %in% big, tissue_plot := "Other"]
  }
  
  # order tissues by best MeanZ
  tis_order <- metrics[, .(mx=max(MeanZ, na.rm=TRUE)), by=tissue_plot][order(-mx)]$tissue_plot
  metrics[, tissue_plot := factor(tissue_plot, levels=tis_order)]
  
  # -----------------------
  # 8) Manhattan x positions grouped by tissue_plot
  # -----------------------
  setorder(metrics, tissue_plot, -MeanZ)
  metrics[, x_in := seq_len(.N), by=tissue_plot]
  block_sizes <- metrics[, .N, by=tissue_plot][order(tissue_plot)]
  block_sizes[, offset := cumsum(shift(N, fill=0))]
  metrics <- merge(metrics, block_sizes[, .(tissue_plot, offset)], by="tissue_plot", all.x=TRUE)
  
  set.seed(1)
  metrics[, x := offset + x_in + runif(.N, -0.25, 0.25)]
  centers <- block_sizes[, .(center = offset + (N+1)/2, tissue_plot)]
  
  block_ext <- copy(block_sizes)
  block_ext[, xmin := offset + 0.5]
  block_ext[, xmax := offset + N + 0.5]
  block_ext[, band := as.integer(as.factor(tissue_plot)) %% 2]
  
  tissue_cols <- if (opt_consistent_tissue_palette) make_tissue_palette(levels(metrics$tissue_plot)) else NULL
  
  # -----------------------
  # 9) Manhattan plot: y=MeanZ ; size=Intensity_used ; label top5 by MeanZ
  # -----------------------
  p1 <- ggplot()
  
  if (opt_block_bg) {
    p1 <- p1 +
      geom_rect(data=block_ext,
                aes(xmin=xmin, xmax=xmax, ymin=-Inf, ymax=Inf, fill=factor(band)),
                alpha=0.08, inherit.aes=FALSE) +
      scale_fill_manual(values=c("0"="grey70","1"="grey20"), guide="none")
  }
  
  top_cells <- metrics[order(-MeanZ)][1:min(top_label_n, nrow(metrics))]
  
  p1 <- p1 +
    geom_point(data = metrics,
             aes(x = x, y = MeanZ, color = tissue_plot, size = Intensity_used),
             alpha = 0.88) +
    scale_x_continuous(breaks=centers$center, labels=centers$tissue_plot, expand=c(0.01, 0.01)) +
    labs(x="Cell lines grouped by tissue",
         y="MeanZ across input regions (higher is better)",
         size=if (opt_intensity_baseline_adjust) "Intensity_adj (median log2 - baseline)" else "Intensity (median log2)") +
    theme_classic(base_size=11) +
    theme(axis.text.x = element_text(angle=45, hjust=1, vjust=1),
          legend.position="right")
  
  if (!is.null(tissue_cols)) p1 <- p1 + scale_color_manual(values=tissue_cols)
  
  if (opt_repel_labels && requireNamespace("ggrepel", quietly=TRUE)) {
    p1 <- p1 + ggrepel::geom_text_repel(
      data=top_cells,
      aes(x=x, y=MeanZ, label=cellline),
      size=2.8, seed=1, min.segment.length=0,
      box.padding=0.25, point.padding=0.25, show.legend=FALSE
    )
  } else {
    p1 <- p1 + geom_text(
      data = top_cells,
      aes(x = x, y = MeanZ, label = cellline),
      size = 2.6,
      vjust = -0.7,
      show.legend = FALSE
    )
  }
  
  if (opt_top_threshold && threshold_type != "none") {
    if (threshold_type == "meanZ_q95") ref_y <- as.numeric(quantile(metrics$MeanZ, 0.95, na.rm=TRUE))
    else if (threshold_type == "top_meanZ_median") ref_y <- median(top_cells$MeanZ, na.rm=TRUE)
    else ref_y <- NA_real_
    if (is.finite(ref_y)) p1 <- p1 + geom_hline(yintercept=ref_y, linetype="dashed", size=0.4, alpha=0.6)
  }
  
  print(p1)
  
}
```

<br>

# Cell-type specificity

```{r celltype_specificity_manhattan, fig.height=9, fig.width=14}
if (!file.exists(accessibility_file)) {
  cat("Specificity file not found: ", accessibility_file, "\n")
  cat("Skipping cell-type specificity section.\n")
} else {
# -----------------------
# Knobs (core)
# -----------------------
min_celllines_per_tissue <- 3
high_quantile <- 0.90
top_label_n <- 5
bar_top_n <- 40

# -----------------------
# Advanced toggles
  # -----------------------
  opt_block_bg      <- TRUE
  opt_top_threshold <- TRUE
  opt_bootstrap_ci  <- FALSE
  
  opt_drop_allzero_regions      <- TRUE
  opt_repel_labels              <- TRUE
  opt_consistent_tissue_palette <- TRUE
  opt_plot_only_big_tissues     <- FALSE
  plot_min_celllines_per_tissue <- 3
  
  opt_intensity_baseline_adjust <- FALSE
  coverage_mode <- "per_cellline"   # "global" | "per_cellline" | "per_tissue" | "global_or_tissue"
  threshold_type <- "meanZ_q95"     # kept name; now applies to NormalizedSumZ
  
  # safety filter
  opt_safety_filter <- TRUE
  min_obs_regions <- 500
  
  # -----------------------
  # Helpers
  # -----------------------
  zscore <- function(x) {
    s <- sd(x, na.rm = TRUE)
    if (!is.finite(s) || s == 0) return(rep(0, length(x)))
    (x - mean(x, na.rm = TRUE)) / s
  }
  
  make_tissue_palette <- function(levels_vec) {
    k <- length(levels_vec)
    if (requireNamespace("Polychrome", quietly = TRUE)) {
      cols <- Polychrome::palette36.colors
      if (k <= length(cols)) return(setNames(cols[seq_len(k)], levels_vec))
    }
    if (requireNamespace("randomcoloR", quietly = TRUE)) {
      cols <- randomcoloR::distinctColorPalette(k)
      return(setNames(cols, levels_vec))
    }
    base <- c(brewer.pal(9, "Set1"), brewer.pal(8, "Dark2"), brewer.pal(8, "Set2"))
    setNames(rep(base, length.out = k), levels_vec)
  }
  
  make_rank_bar <- function(df, value_col, title, higher_better = TRUE, top_n = 40) {
    d <- copy(df)
    if (higher_better) d <- d[order(-get(value_col))] else d <- d[order(get(value_col))]
    d <- d[1:min(top_n, nrow(d))]
    d[, cellline := factor(cellline, levels = rev(cellline))]
    ggplot(d, aes(x = cellline, y = .data[[value_col]], fill = tissue_plot)) +
      geom_col(width = 0.85) +
      coord_flip() +
      labs(x = NULL, y = title) +
      theme_classic(base_size = 11) +
      theme(panel.grid.major.y = element_blank(),
            panel.grid.minor = element_blank(),
            legend.position = "right")
  }
  
  write_mat_tsv <- function(mat, file) {
    dt_out <- as.data.table(mat, keep.rownames = "region")
    fwrite(dt_out, file = file, sep = "\t", quote = FALSE)
  }
  
  top2_info <- function(x_named) {
    o <- order(x_named, decreasing = TRUE)
    list(
      winner = names(x_named)[o[1]],
      winner_val = as.numeric(x_named[o[1]]),
      runner = names(x_named)[o[2]],
      runner_val = as.numeric(x_named[o[2]])
    )
  }
  
  # -----------------------
  # 1) Read activity; collapse duplicates by MAX
  # -----------------------
  dt_all <- fread(accessibility_file, header = FALSE, sep = "\t", col.names = c("region", "cellline", "activity"))
  dt_all <- dt_all[, .(activity = max(activity, na.rm = TRUE)), by = .(region, cellline)]
  
  cell <- fread(tissue_map_sc, header = FALSE, sep = "\t", col.names = c("cellline", "tissue"))
  cell <- cell[!is.na(cellline) & cellline != "" & !is.na(tissue) & tissue != ""]
  
  enh_count <- fread(enhancer_count_file, header = FALSE, sep = "\t", col.names = c("cellline", "enhancer_count"))
  enh_count <- enh_count[!is.na(cellline) & cellline != "" & !is.na(enhancer_count)]
  
  # restrict to mapped cell lines
  dt_all <- dt_all[cell, on = "cellline", nomatch = 0]
  
  # keep tissues with >= N cell lines
  tab_tis <- table(cell$tissue)
  keep_tis <- names(tab_tis[tab_tis >= min_celllines_per_tissue])
  cell <- cell[tissue %in% keep_tis]
  dt_all <- dt_all[cell, on = "cellline", nomatch = 0]
  
  # optional: restrict to input regions file
  dt <- copy(dt_all)
  if (!is.na(regions_file)) {
    rr <- fread(regions_file, header = FALSE)
    input_regions <- unique(rr[[1]])
    dt <- dt[region %in% input_regions]
#    message("Using regions_file: N=", length(input_regions), " | found in table: N=", length(unique(dt$region)))
  } else {
#    message("No regions_file: using regions present in accessibility_file: N=", length(unique(dt$region)))
  }
  
  # -----------------------
  # 2) Pseudocount
  # -----------------------
  pc <- suppressWarnings(min(dt_all$activity[dt_all$activity > 0]))
  if (!is.finite(pc) || pc <= 0) pc <- 1e-12
  
  # -----------------------
  # 3) Build FULL matrix region x cellline (fill missing as 0)
  # -----------------------
  mm_dt <- dcast(dt, region ~ cellline, value.var = "activity", fill = 0)
  regions_u <- mm_dt$region
  mm_dt[, region := NULL]
  mm <- as.matrix(mm_dt)
  rownames(mm) <- regions_u
  rm(mm_dt)
  
  cell_u <- unique(cell[, .(cellline, tissue)])
  cell_u <- cell_u[cellline %in% colnames(mm)]
  setorder(cell_u, tissue, cellline)
  mm <- mm[, cell_u$cellline, drop = FALSE]
  
  # drop all-zero regions across included cell lines
  if (opt_drop_allzero_regions) {
    keep_r <- which(rowSums(mm > 0) > 0)
    mm <- mm[keep_r, , drop = FALSE]
#    message("Dropped all-zero regions: kept ", nrow(mm), " regions.")
  }
  
  # per cell line: how many regions have non-zero activity
  n_active_regions <- colSums(mm > 0, na.rm = TRUE)
  
  # safety filter
  if (opt_safety_filter) {
    keep_cl <- names(n_active_regions)[n_active_regions >= min_obs_regions]
    if (length(keep_cl) < 2) stop("Too few cell lines remain after safety filter. Lower min_obs_regions.")
    dropped_cl <- setdiff(colnames(mm), keep_cl)
    if (length(dropped_cl) > 0) {
#      message("Safety filter dropped ", length(dropped_cl), " cell lines with n_obs_regions < ", min_obs_regions)
    }
    mm <- mm[, keep_cl, drop = FALSE]
    cell_u <- cell_u[cellline %in% keep_cl]
    setorder(cell_u, tissue, cellline)
    n_active_regions <- colSums(mm > 0, na.rm = TRUE)
  }
  
  logM <- log2(mm + pc)
  
  # -----------------------
  # 4) Baseline (optional)
  # -----------------------
  BaselineIntensity <- NULL
  if (opt_intensity_baseline_adjust) {
    mm_all_dt <- dcast(dt_all, region ~ cellline, value.var = "activity", fill = 0)
    mm_all_dt[, region := NULL]
    mm_all <- as.matrix(mm_all_dt)
    mm_all <- mm_all[, intersect(colnames(mm_all), cell_u$cellline), drop = FALSE]
    log_all <- log2(mm_all + pc)
    BaselineIntensity <- apply(log_all, 2, median, na.rm = TRUE)
    rm(mm_all_dt, mm_all, log_all)
  }
  
  # -----------------------
  # 5) Coverage thresholds
  # -----------------------
  mm_all_for_thr <- dcast(dt_all, region ~ cellline, value.var = "activity", fill = 0)
  mm_all_for_thr[, region := NULL]
  log_all_vals <- as.vector(log2(as.matrix(mm_all_for_thr) + pc))
  thr_high_global <- as.numeric(quantile(log_all_vals, high_quantile, na.rm = TRUE))
  rm(mm_all_for_thr, log_all_vals)
  
  thr_cell <- NULL
  if (coverage_mode == "per_cellline") {
    mm_all2 <- dcast(dt_all, region ~ cellline, value.var = "activity", fill = 0)
    mm_all2[, region := NULL]
    log_all2 <- log2(as.matrix(mm_all2) + pc)
    colnames(log_all2) <- colnames(mm_all2)
    thr_cell <- apply(log_all2, 2, function(x) as.numeric(quantile(x, high_quantile, na.rm = TRUE)))
    rm(mm_all2, log_all2)
  }
  
  thr_tis <- NULL
  if (coverage_mode %in% c("per_tissue", "global_or_tissue")) {
    mm_all3 <- dcast(dt_all, region ~ cellline, value.var = "activity", fill = 0)
    mm_all3[, region := NULL]
    log_all3 <- log2(as.matrix(mm_all3) + pc)
    colnames(log_all3) <- colnames(mm_all3)
    col_tissue_map <- cell_u[match(colnames(log_all3), cell_u$cellline), tissue]
    thr_tis <- sapply(unique(col_tissue_map), function(tis) {
      cols <- which(col_tissue_map == tis)
      as.numeric(quantile(as.vector(log_all3[, cols, drop = FALSE]), high_quantile, na.rm = TRUE))
    })
    rm(mm_all3, log_all3, col_tissue_map)
  }
  
  # -----------------------
  # 6) Metrics computed on FULL matrix logM
  # -----------------------
  Intensity <- apply(logM, 2, median, na.rm = TRUE)
  # Intensity_q <- 0.75
  # Intensity <- apply(logM, 2, function(x) as.numeric(quantile(x, probs = Intensity_q, na.rm = TRUE)))
  
  if (coverage_mode == "global") {
    Coverage <- apply(logM, 2, function(x) mean(x >= thr_high_global, na.rm = TRUE))
  } else if (coverage_mode == "per_cellline") {
    Coverage <- mapply(function(col, thr) mean(col >= thr, na.rm = TRUE),
                       as.data.frame(logM), thr_cell[colnames(logM)])
    Coverage <- as.numeric(Coverage)
    names(Coverage) <- colnames(logM)
  } else if (coverage_mode == "per_tissue") {
    Coverage <- sapply(colnames(logM), function(cl) {
      tis <- cell_u[cellline == cl, tissue][1]
      mean(logM[, cl] >= thr_tis[[tis]], na.rm = TRUE)
    })
  } else if (coverage_mode == "global_or_tissue") {
    Coverage <- sapply(colnames(logM), function(cl) {
      tis <- cell_u[cellline == cl, tissue][1]
      thr <- max(thr_high_global, thr_tis[[tis]])
      mean(logM[, cl] >= thr, na.rm = TRUE)
    })
  } else stop("Unknown coverage_mode: ", coverage_mode)
  
  rankM <- t(apply(logM, 1, function(x) rank(-x, ties.method = "average")))
  MeanRank <- colMeans(rankM, na.rm = TRUE)
  
  rowMean <- rowMeans(logM, na.rm = TRUE)
  rowSd <- apply(logM, 1, sd, na.rm = TRUE)
  rowSd[rowSd == 0 | !is.finite(rowSd)] <- NA_real_
  
  zM <- (logM - rowMean) / rowSd
  zM[!is.finite(zM)] <- NA_real_
  
  SumZ <- colSums(zM, na.rm = TRUE)
  MeanZ <- SumZ / nrow(zM)
  
  # normalize SumZ by enhancer count
  enh_count_vec <- enh_count$enhancer_count
  names(enh_count_vec) <- enh_count$cellline
  denom <- enh_count_vec[names(SumZ)]
  denom[is.na(denom) | denom <= 0] <- NA_real_
  NormalizedSumZ <- SumZ / denom
  
#  write_mat_tsv(zM, out_z_mat)
#  message("Saved: ", out_z_mat)
  
  # -----------------------
  # assemble table
  # -----------------------
  metrics <- data.table(
    cellline = cell_u$cellline,
    tissue   = cell_u$tissue
  )
  metrics[, n_active_regions := as.integer(n_active_regions[cellline])]
  metrics[, n_obs_regions := n_active_regions]
  metrics[, frac_active_regions := n_active_regions / nrow(mm)]
  metrics[, n_regions_total := nrow(mm)]
  metrics[, Intensity := as.numeric(Intensity[cellline])]
  metrics[, Coverage := as.numeric(Coverage[cellline])]
  metrics[, MeanRank := as.numeric(MeanRank[cellline])]
  metrics[, SumZ := as.numeric(SumZ[cellline])]
  metrics[, MeanZ := as.numeric(MeanZ[cellline])]
  metrics[, enhancer_count := as.numeric(enh_count_vec[cellline])]
  metrics[, NormalizedSumZ := as.numeric(NormalizedSumZ[cellline])]
  metrics[, n_regions := nrow(mm)]
  
  if (opt_intensity_baseline_adjust) {
    metrics[, BaselineIntensity := as.numeric(BaselineIntensity[cellline])]
    metrics[, Intensity_adj := Intensity - BaselineIntensity]
    metrics[, Intensity_used := Intensity_adj]
  } else {
    metrics[, BaselineIntensity := NA_real_]
    metrics[, Intensity_adj := NA_real_]
    metrics[, Intensity_used := Intensity]
  }
  
  metrics[, Coverage_used := Coverage]
  
  metrics[, ZI := zscore(Intensity_used)]
  metrics[, ZC := zscore(Coverage_used)]
  metrics[, ZR := zscore(-MeanRank)]
  metrics[, ZZ := zscore(SumZ)]
  metrics[, OriginScore := ZI + ZC + ZR + ZZ]
  
#  fwrite(metrics[order(-NormalizedSumZ)], file = out_metrics_table, sep = "\t", quote = FALSE)
#  message("Saved metrics table: ", out_metrics_table)
  
  # -----------------------
  # 7) Plot-only tissue grouping
  # -----------------------
  metrics[, tissue_plot := as.character(tissue)]
  if (opt_plot_only_big_tissues) {
    tc <- metrics[, .N, by = tissue_plot]
    big <- tc[N >= plot_min_celllines_per_tissue, tissue_plot]
    metrics[!tissue_plot %in% big, tissue_plot := "Other"]
  }
  
  # order tissues by best NormalizedSumZ
  tis_order <- metrics[, .(mx = max(NormalizedSumZ, na.rm = TRUE)), by = tissue_plot][order(-mx)]$tissue_plot
  metrics[, tissue_plot := factor(tissue_plot, levels = tis_order)]
  
  # -----------------------
  # 8) Manhattan x positions grouped by tissue_plot
  # -----------------------
  setorder(metrics, tissue_plot, -NormalizedSumZ)
  metrics[, x_in := seq_len(.N), by = tissue_plot]
  block_sizes <- metrics[, .N, by = tissue_plot][order(tissue_plot)]
  block_sizes[, offset := cumsum(shift(N, fill = 0))]
  metrics <- merge(metrics, block_sizes[, .(tissue_plot, offset)], by = "tissue_plot", all.x = TRUE)
  
  set.seed(1)
  metrics[, x := offset + x_in + runif(.N, -0.25, 0.25)]
  centers <- block_sizes[, .(center = offset + (N + 1) / 2, tissue_plot)]
  
  block_ext <- copy(block_sizes)
  block_ext[, xmin := offset + 0.5]
  block_ext[, xmax := offset + N + 0.5]
  block_ext[, band := as.integer(as.factor(tissue_plot)) %% 2]
  
  tissue_cols <- if (opt_consistent_tissue_palette) make_tissue_palette(levels(metrics$tissue_plot)) else NULL
  
  # -----------------------
  # 9) Manhattan plot: y = NormalizedSumZ ; size = Intensity_used ; label top5
  # -----------------------
  p1 <- ggplot()
  
  if (opt_block_bg) {
    p1 <- p1 +
      geom_rect(data = block_ext,
                aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf, fill = factor(band)),
                alpha = 0.08, inherit.aes = FALSE) +
      scale_fill_manual(values = c("0" = "grey70", "1" = "grey20"), guide = "none")
  }
  
  top_cells <- metrics[order(-NormalizedSumZ)][1:min(top_label_n, nrow(metrics))]
  
  p1 <- p1 +
    geom_point(data = metrics,
               aes(x = x, y = NormalizedSumZ, color = tissue_plot, size = Intensity_used),
               alpha = 0.88) +
    scale_x_continuous(breaks = centers$center, labels = centers$tissue_plot, expand = c(0.01, 0.01)) +
    labs(x = "Cell lines grouped by tissue",
         y = "Normalized SumZ = SumZ / enhancer_count",
         size = if (opt_intensity_baseline_adjust) "Intensity_adj (median log2 - baseline)" else "Intensity (median log2)") +
    theme_classic(base_size = 11) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
          legend.position = "right")
  
  if (!is.null(tissue_cols)) p1 <- p1 + scale_color_manual(values = tissue_cols)
  
  if (opt_repel_labels && requireNamespace("ggrepel", quietly = TRUE)) {
    p1 <- p1 + ggrepel::geom_text_repel(
      data = top_cells,
      aes(x = x, y = NormalizedSumZ, label = cellline),
      size = 2.8, seed = 1, min.segment.length = 0,
      box.padding = 0.25, point.padding = 0.25, show.legend = FALSE
    )
  } else {
    p1 <- p1 + geom_text(
      data = top_cells,
      aes(x = x, y = NormalizedSumZ, label = cellline),
      size = 2.6,
      vjust = -0.7,
      show.legend = FALSE
    )
  }
  
  if (opt_top_threshold && threshold_type != "none") {
    if (threshold_type == "meanZ_q95") {
      ref_y <- as.numeric(quantile(metrics$NormalizedSumZ, 0.95, na.rm = TRUE))
    } else if (threshold_type == "top_meanZ_median") {
      ref_y <- median(top_cells$NormalizedSumZ, na.rm = TRUE)
    } else ref_y <- NA_real_
    if (is.finite(ref_y)) p1 <- p1 + geom_hline(yintercept = ref_y, linetype = "dashed", size = 0.4, alpha = 0.6)
  }
  
  print(p1)
  
  # optional cleanup
  rm(rankM, zM, rowMean, rowSd, mm, logM)
}

```

<br>

# TF binding enrichment

```{r tf_binding_enrichment, fig.height=10, fig.width=10}
colors <- colorRampPalette(brewer.pal(9, "Oranges"))(250)
colors_A <- colorRampPalette(brewer.pal(9, "Greens"))(250)
colors_B <- colorRampPalette(brewer.pal(9, "Blues"))(250)
col_breaks <- seq(0, 1, length=251)

if (!file.exists(m_tf_path)) {
  cat("Missing TF enrichment file: ", m_tf_path, "\n")
} else if (!file.exists(tfbs_summary)) {
  cat("Missing TFBS summary file: ", tfbs_summary, "\n")
} else {
  m_tf <- read.table(m_tf_path, sep="\t", header=FALSE, quote="", as.is=TRUE)
  s_tf <- read.table(tfbs_summary, sep="\t", header=FALSE, quote="", as.is=TRUE)

  m_tf_p  <- m_tf[m_tf$V1 =="promoter-TSS",]
  m_tf_in <- m_tf[m_tf$V1 =="intron",]
  m_tf_it <- m_tf[m_tf$V1 =="intergenic",]

  n <- nrow(file)

  m_tf$V5 <- ave(m_tf$V4, m_tf[,c(2,3)], FUN = sum)
  m_tf <- unique(m_tf[,c(2,3,5)])

  if(nrow(m_tf) > 2){
    for (x in 1:nrow(m_tf)) {
      temp <- s_tf[(s_tf$V1==m_tf[x,1]) & (s_tf$V2==m_tf[x,2]), 3]
      m_tf[x,"jac"] <- (m_tf[x,3])/(temp+n-m_tf[x,3])
      m_tf[x,"A"]   <- (m_tf[x,3])/(temp)
      m_tf[x,"B"]   <- (m_tf[x,3])/(n)
    }

    top_tf <- head(m_tf[order(m_tf$jac, decreasing = TRUE),], 100)

    n_row <- length(unique(top_tf$V3))
    n_col <- length(unique(top_tf$V2))
    mm_tf <- matrix(NA_real_, n_row, n_col)
    colnames(mm_tf) <- unique(top_tf$V2)
    rownames(mm_tf) <- unique(top_tf$V3)

    for (x in 1:nrow(m_tf)) {
      if((m_tf[x,1] %in% colnames(mm_tf)) & (m_tf[x,2] %in% rownames(mm_tf)))
        mm_tf[m_tf[x,2], m_tf[x,1]] <- m_tf[x,"jac"]
    }

    temp1 <- apply(mm_tf,1,mean,na.rm=TRUE)
    mm_tf <- mm_tf[order(-temp1),,drop=FALSE]
    temp2 <- apply(mm_tf, 2, function(x){sum(!is.na(x))})
    mm_tf <- t(mm_tf); mm_tf <- mm_tf[order(-temp2),,drop=FALSE]; mm_tf <- t(mm_tf)

    pheatmap(mm_tf, main="Overall TFs enrichment", na_col="gray",
             cluster_rows=FALSE, cluster_cols=FALSE, color=colors, breaks=col_breaks)
  }
}
```

<br>

# Functional enrichment of target genes

```{r webgestalt_load, include=FALSE}
has_webgestalt <- requireNamespace("WebGestaltR", quietly = TRUE)
```

## KEGG

```{r kegg_enrichment, results='hide'}
if (!has_webgestalt) {
  cat("WebGestaltR not installed; skipping gene enrichment sections.\n")
} else if (!file.exists(gene_path)) {
  cat("Missing target gene file: ", gene_path, "\n")
} else {

  library(WebGestaltR)

  gene <- read.table(gene_path, sep="\t", header=FALSE, quote="", fill=TRUE, as.is=TRUE)
  outputDirectory <- outdir

  gene1 <- gene[1,]; gene1 <- gene1[gene1 != "" & gene1 != "NA"]
  gene2 <- gene[2,]; gene2 <- gene2[gene2 != "" & gene2 != "NA"]
  gene3 <- gene[3,]; gene3 <- gene3[gene3 != "" & gene3 != "NA"]
  gene4 <- gene[4,]; gene4 <- gene4[gene4 != "" & gene4 != "NA"]
  gene5 <- gene[5,]; gene5 <- gene5[gene5 != "" & gene5 != "NA"]

  enrichResult1 <- WebGestaltR(enrichMethod="ORA", organism="hsapiens", enrichDatabase="pathway_KEGG",
                              interestGene=gene1, interestGeneType="genesymbol",
                              referenceSet="genome", referenceGeneType="genesymbol",
                              isOutput=FALSE, outputDirectory=outputDirectory, projectName=NULL)
  enrichResult2 <- WebGestaltR(enrichMethod="ORA", organism="hsapiens", enrichDatabase="pathway_KEGG",
                              interestGene=gene2, interestGeneType="genesymbol",
                              referenceSet="genome", referenceGeneType="genesymbol",
                              isOutput=FALSE, outputDirectory=outputDirectory, projectName=NULL)
  enrichResult3 <- WebGestaltR(enrichMethod="ORA", organism="hsapiens", enrichDatabase="pathway_KEGG",
                              interestGene=gene3, interestGeneType="genesymbol",
                              referenceSet="genome", referenceGeneType="genesymbol",
                              isOutput=FALSE, outputDirectory=outputDirectory, projectName=NULL)
  enrichResult4 <- WebGestaltR(enrichMethod="ORA", organism="hsapiens", enrichDatabase="pathway_KEGG",
                              interestGene=gene4, interestGeneType="genesymbol",
                              referenceSet="genome", referenceGeneType="genesymbol",
                              isOutput=FALSE, outputDirectory=outputDirectory, projectName=NULL)
  enrichResult5 <- WebGestaltR(enrichMethod="ORA", organism="hsapiens", enrichDatabase="pathway_KEGG",
                              interestGene=gene5, interestGeneType="genesymbol",
                              referenceSet="genome", referenceGeneType="genesymbol",
                              isOutput=FALSE, outputDirectory=outputDirectory, projectName=NULL)
}
```

```{r kegg_top_dotplot, fig.show="hold", out.width="90%"}
if (has_webgestalt && exists("enrichResult1")) {
  g <- NULL
  if(!is.null(enrichResult1)){ g1 <- head(enrichResult1[order(enrichResult1$FDR),c(2,7,9)],10); g1$source <- "closest"; g <- rbind(g,g1) }
  if(!is.null(enrichResult2)){ g2 <- head(enrichResult2[order(enrichResult2$FDR),c(2,7,9)],10); g2$source <- "genes in 50kb"; g <- rbind(g,g2) }
  if(!is.null(enrichResult3)){ g3 <- head(enrichResult3[order(enrichResult3$FDR),c(2,7,9)],10); g3$source <- "FANTOM5"; g <- rbind(g,g3) }
  if(!is.null(enrichResult4)){ g4 <- head(enrichResult4[order(enrichResult4$FDR),c(2,7,9)],10); g4$source <- "validated"; g <- rbind(g,g4) }
  if(!is.null(enrichResult5)){ g5 <- head(enrichResult5[order(enrichResult5$FDR),c(2,7,9)],10); g5$source <- "eQTL"; g <- rbind(g,g5) }

  if(!is.null(g) && nrow(g) > 0){
    g$adjFDR <- -log10(g$FDR)
    g$source <- factor(g$source,levels=c("closest","genes in 50kb","validated","FANTOM5","eQTL"))
    g$count <- 1
    g$support <- ave(g$count,g$description,FUN=sum)
    g$sumR <- ave(g$enrichmentRatio,g$description,FUN=sum)
    g <- g[order(g$support,g$sumR),]
    temp <- unique(g$description)
    g$description <- factor(g$description, levels=temp)

    ggplot(g, aes(x=adjFDR, y=description, size=enrichmentRatio, color=source)) +
      geom_point(alpha=0.8) +
      theme_classic() +
      scale_size(range=c(2,6)) +
      xlab("-log10 FDR") + ylab("")
  } else {
    cat("No KEGG enrichment results.\n")
  }
}
```

```{r kegg_full_tables, echo=FALSE, results='asis'}
if (has_webgestalt && exists("enrichResult1")) {
  if(!is.null(enrichResult1)){
    DT::datatable(enrichResult1[order(enrichResult1$FDR),c(1,2,4:9)], rownames=FALSE,
                  caption="functional enrichment of closest gene",
                  extensions="Buttons",
                  options=list(dom="Bfrtip", buttons=c("excel","csv")))
  }
  if(!is.null(enrichResult2)){
    DT::datatable(enrichResult2[order(enrichResult2$FDR),c(1,2,4:9)], rownames=FALSE,
                  caption="functional enrichment of genes within a distance of 50kb",
                  extensions="Buttons",
                  options=list(dom="Bfrtip", buttons=c("excel","csv")))
  }
  if(!is.null(enrichResult3)){
    DT::datatable(enrichResult3[order(enrichResult3$FDR),c(1,2,4:9)], rownames=FALSE,
                  caption="functional enrichment of target genes in FANTOM5",
                  extensions="Buttons",
                  options=list(dom="Bfrtip", buttons=c("excel","csv")))
  }
  if(!is.null(enrichResult4)){
    DT::datatable(enrichResult4[order(enrichResult4$FDR),c(1,2,4:9)], rownames=FALSE,
                  caption="functional enrichment of validated target genes",
                  extensions="Buttons",
                  options=list(dom="Bfrtip", buttons=c("excel","csv")))
  }
  if(!is.null(enrichResult5)){
    DT::datatable(enrichResult5[order(enrichResult5$FDR),c(1,2,4:9)], rownames=FALSE,
                  caption="functional enrichment of target genes defined by eQTL",
                  extensions="Buttons",
                  options=list(dom="Bfrtip", buttons=c("excel","csv")))
  }
}
```

<br>

# GWAS SNPs enrichment

```{r gwas_enrichment, fig.height=10, fig.width=10}
if (!file.exists(m_snp_path)) {
  cat("Missing GWAS enrichment file: ", m_snp_path, "\n")
} else if (!file.exists(snp_summary)) {
  cat("Missing GWAS summary file: ", snp_summary, "\n")
} else {
  m_snp <- read.table(m_snp_path, sep="\t", header=FALSE, quote="", as.is=TRUE)
  s_snp <- read.table(snp_summary, sep="\t", header=FALSE, quote="", as.is=TRUE)

  if (nrow(m_snp) > 2) {
    m_snp$V3 <- 0
    for (x in 1:nrow(m_snp)) {
      temp <- s_snp[s_snp$V1==m_snp[x,1],2]
      m_snp[x,3] <- (m_snp[x,2])/(temp+n-m_snp[x,2])
    }
    names(m_snp) <- c("Trait","Overlap","Jaccard Index")
    DT::datatable(
      m_snp[order(m_snp$`Jaccard Index`, decreasing = TRUE), c(1,3)],
      rownames=FALSE,
      caption="GWAS SNP enrichment",
      extensions="Buttons",
      options=list(dom="Bfrtip", buttons=c("excel","csv"))
    ) %>% DT::formatRound(columns=2, digits=5)
  } else {
    cat("GWAS enrichment table is empty or too small.\n")
  }
}
```

<br>

# GTEx variants enrichment

```{r gtex_eqtl_enrichment, fig.height=10, fig.width=10}
if (!file.exists(m_eqtl_path)) {
  cat("Missing eQTL enrichment file: ", m_eqtl_path, "\n")
} else if (!file.exists(eqtl_summary)) {
  cat("Missing eQTL summary file: ", eqtl_summary, "\n")
} else {
  m_eqtl <- read.table(m_eqtl_path, sep="\t", header=FALSE, quote="", as.is=TRUE)
  s_eqtl <- read.table(eqtl_summary, sep="\t", header=FALSE, quote="", as.is=TRUE)

  if (nrow(m_eqtl) > 2) {
    m_eqtl$V3 <- 0
    for (x in 1:nrow(m_eqtl)) {
      temp <- s_eqtl[s_eqtl$V1==m_eqtl[x,1],2]
      m_eqtl[x,3] <- (m_eqtl[x,2])/(temp+n-m_eqtl[x,2])
    }
    names(m_eqtl) <- c("Tissue","Overlap","Jaccard Index")
    DT::datatable(
      m_eqtl[order(m_eqtl$`Jaccard Index`, decreasing = TRUE), c(1,3)],
      rownames=FALSE,
      caption="eQTL variant enrichment",
      extensions="Buttons",
      options=list(dom="Bfrtip", buttons=c("excel","csv"))
    ) %>% DT::formatRound(columns=2, digits=5)
  } else {
    cat("eQTL enrichment table is empty or too small.\n")
  }
}
```
